Generation of TypeScript Declaration Files from JavaScript Code

### 2020-08-26 programming-2021/main.tex
Add table with dts-parse statistics

We implemented only basic TypeScript features and we focused our effort in building an
end-to-end solution that generates valid and useful declaration files. Only the basic
types \texttt{string}, \texttt{number}, \texttt{boolean} and its union were considered
both for function parameters and interface properties. Common types such as
\texttt{any}, arrays, function callbacks, tuples, intersections and features such as
generics, index signatures were not implemented. Each declaration file was tagged with
\texttt{dts-parse} and those containing unimplemented features were ignored.

@@

\paragraph*{Further types}
Similar approaches could be conceived for indexed types, polymorphic types, dependent
types, etc. However, at present we believe incorporating these features in a tool to be
overblown because they are not widely used in DefinitelyTyped (see
Table~\ref{table-of-occurrences}).

>>>>

We implemented only basic TypeScript features and we focused our effort in building an
end-to-end solution that generates valid and useful declaration files. Only the basic
types \texttt{string}, \texttt{number}, \texttt{boolean}, its union and the optional type were considered
both for function parameters and interface properties. Common types such as
\texttt{any}, arrays, function callbacks, tuples, intersections and features such as
generics, index signatures were not implemented.

Each declaration file in the DT repository was tagged with
\texttt{dts-parse}. We summarized the results in Table~\ref{tab:dts-parse-stats}. We based our decisions of prioritizing particular features on these results.

\begin{table}[tp]
  \begin{center}
    \begin{tabular}{ |c|c|c| } 
      \hline
      \textbf{TypeScript feature} & \textbf{Count Declaration Files} & \textbf{\%} \\ 
      \hline
      string & 5086 & 76.50\% \\
      optional-parameter & 4915 & 73.93\% \\
      boolean & 3891 & 58.53\% \\
      number & 3699 & 55.64\% \\
      void & 3548 & 53.37\% \\
      type-union & 3456 & 51.99\% \\
      type-function & 3286 & 49.43\% \\
      type-array & 3264 & 49.10\% \\
      type-any & 3127 & 47.04\% \\
      literals & 1925 & 28.96\% \\
      alias-type & 1899 & 28.56\% \\
      index-signature & 1271 & 19.12\% \\
      generics-function & 1145 & 17.22\% \\
      dot-dot-dot-token & 882 & 13.27\% \\
      call-signature & 817 & 12.29\% \\
      generics-interface & 718 & 10.80\% \\
      object & 661 & 9.94\% \\
      undefined & 577 & 8.68\% \\
      type-intersection & 431 & 6.48\% \\
      readonly & 373 & 5.61\% \\
      type-tuple & 355 & 5.34\% \\
      generics-class & 265 & 3.99\% \\
      static & 251 & 3.78\% \\
      private & 82 & 1.23\% \\
      public & 43 & 0.65\% \\
      protected & 19 & 0.29\% \\
      \hline
    \end{tabular}
  \end{center}
  \caption{Distribution of TypeScript features in declaration files uploaded to DefinitelyTyped. The 2nd column shows how many files in DT make use of a specific TypeScript feature. The 3rd column shows the corresponding percentage for a total of 6648 declaration files in DT. Numbers were obtained from commit \texttt{da93d7b13094bacb170ead3f4a289f3b8687e4e5} (April 4, 2020).}
  \label{tab:dts-parse-stats}
\end{table}

@@

\paragraph*{Further types}
Similar approaches could be conceived for indexed types, polymorphic types, dependent
types, etc. However, at present we believe incorporating these features in a tool to be
overblown because they are not widely used in DefinitelyTyped (see
Table~\ref{tab:dts-parse-stats}).
